---
title: "Statistical Analysis"
author: "Francesco Pisu"
date: '2022-04-13'
output: html_document
---

```{r}
library("DescTools")
```


```{r}
train = read.csv('../input/train.csv')
external = read.csv('../input/external.csv')
```


# Derivation cohort
```{r}
s1 <- train[1:123, ]
s2 <- train[201:240, ]
pop <- rbind(s1, s2)

pos <- pop[pop$symptoms == 1,]
neg <- pop[pop$symptoms == 0,]

postrain <- train[train$symptoms == 1,]
negtrain <- train[train$symptoms == 0,]

pvalues <- list()
```

```{r}
MedianCI(pop[, "age"], R = 2000)
```


## Binary Variables

### Age
```{r}
varname <- "age"
MedianCI(pos[, varname], method="boot", R = 2000)
MedianCI(neg[, varname], method="boot", R = 2000)

r <- wilcox.test(pos[, varname], neg[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### Gender
```{r}
varname <- "gender"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)))
r
pvalues <- append(pvalues, r$p.value)
```

### Hypertension
```{r}
varname <- "hypertension"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)))
r
pvalues <- append(pvalues, r$p.value)
```

### CAD
```{r}
varname <- "CAD"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)))
r
pvalues <- append(pvalues, r$p.value)
```

### Smoker
```{r}
varname <- "smoker"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)))
r
pvalues <- append(pvalues, r$p.value)
```

### Diabetes
```{r}
varname <- "diabetes"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)),
               )
r
pvalues <- append(pvalues, r$p.value)
```


### Dyslipidemia
```{r}
varname <- "lipids"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)))
r
pvalues <- append(pvalues, r$p.value)
```

### Stenosis
```{r}
varname <- "stenosis"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)))
r
pvalues <- append(pvalues, r$p.value)
```

### Presence of IPH
```{r}
varname <- "iph"

table(train[, varname])

table(postrain[, varname])
prop.table(table(postrain[, varname]))

table(negtrain[, varname])
prop.table(table(negtrain[, varname]))

posf <- postrain[postrain[, varname] == 1,]
negf <- negtrain[negtrain[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(postrain), nrow(negtrain)))
r
pvalues <- append(pvalues, r$p.value)
```


## Continuous Variables
```{r}
install.packages("olsrr")
install.packages("DescTools")
library("olsrr")
library("ggplot2")
library("DescTools")
```

### Total Plaque Volume
```{r}
varname <- "plaque_volume"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### Lipid Volume
```{r}
varname <- "lipid_volume"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### Mixed Volume
```{r}
varname <- "mixed_volume"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```
### Calcium Volume
```{r}
varname <- "calcium_volume"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### IPH Volume
```{r}
varname <- "iph_volume"
MedianCI(postrain[, varname], method="exact", R = 2000)
MedianCI(negtrain[, varname], method="exact", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

#pvalues <- append(pvalues, r$p.value)
```

### Lipid - IPH Volume
```{r}
varname <- "lipid_minus_iph"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### % Lipid volume
```{r}
varname <- "perc_lipid"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### % Mixed volume
```{r}
varname <- "perc_mixed"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### % Calcium volume
```{r}
varname <- "perc_calcium"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### % IPH volume
```{r}
varname <- "perc_iph"
MedianCI(postrain[, varname], method="exact", R = 2000)
MedianCI(negtrain[, varname], method="exact", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```



### % Lipid - IPH volume
```{r}
varname <- "perc_lipid_minus_iph"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```


### IPH to Lipid volume ratio
```{r}
varname <- "iph_total_lipid_ratio"
MedianCI(postrain[, varname], method="exact", R = 2000)
MedianCI(negtrain[, varname], method="exact", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

```{r}
p.adjust(pvalues, method="BH")
```


# External Validation Cohort
```{r}
pos <- external[external$symptoms == 1,]
neg <- external[external$symptoms == 0,]

pop <- external

postrain <- external[external$symptoms == 1,]
negtrain <- external[external$symptoms == 0,]

pvalues <- list()
```

## Binary Variables

### Age
```{r}
varname <- "age"
MedianCI(pos[, varname], method="boot", R = 2000)
MedianCI(neg[, varname], method="boot", R = 2000)

r <- wilcox.test(pos[, varname], neg[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### Gender
```{r}
varname <- "gender"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)))
r
pvalues <- append(pvalues, r$p.value)
```

### Hypertension
```{r}
varname <- "hypertension"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)))
r
pvalues <- append(pvalues, r$p.value)
```

### CAD
```{r}
varname <- "CAD"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)))
r
pvalues <- append(pvalues, r$p.value)
```

### Smoker
```{r}
varname <- "smoker"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)))
r
pvalues <- append(pvalues, r$p.value)
```

### Diabetes
```{r}
varname <- "diabetes"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)),
               )
r
pvalues <- append(pvalues, r$p.value)
```


### Dyslipidemia
```{r}
varname <- "lipids"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)))
r
pvalues <- append(pvalues, r$p.value)
```
```{r}
varname <- "symptoms"

table(train[, varname])
```

### Stenosis
```{r}
varname <- "stenosis"

table(pop[, varname])

table(pos[, varname])
prop.table(table(pos[, varname]))

table(neg[, varname])
prop.table(table(neg[, varname]))

posf <- pos[pos[, varname] == 1,]
negf <- neg[neg[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(pos), nrow(neg)))
r
pvalues <- append(pvalues, r$p.value)
```

### Presence of IPH
```{r}
varname <- "iph"

table(pop[, varname])

table(postrain[, varname])
prop.table(table(postrain[, varname]))

table(negtrain[, varname])
prop.table(table(negtrain[, varname]))

posf <- postrain[postrain[, varname] == 1,]
negf <- negtrain[negtrain[, varname] == 1,]

r <- prop.test(c(nrow(posf), nrow(negf)), c(nrow(postrain), nrow(negtrain)))
r
#pvalues <- append(pvalues, r$p.value)
```


## Continuous Variables

### Total Plaque Volume
```{r}
varname <- "plaque_volume"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### Lipid Volume
```{r}
varname <- "lipid_volume"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### Mixed Volume
```{r}
varname <- "mixed_volume"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```
### Calcium Volume
```{r}
varname <- "calcium_volume"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### IPH Volume
```{r}
varname <- "iph_volume"
MedianCI(postrain[, varname], method="exact")
MedianCI(negtrain[, varname], method="exact")

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```


### Lipid - IPH Volume
```{r}
varname <- "lipid_minus_iph"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### % Lipid volume
```{r}
varname <- "perc_lipid"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### % Mixed volume
```{r}
varname <- "perc_mixed"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### % Calcium volume
```{r}
varname <- "perc_calcium"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

### % IPH volume
```{r}
varname <- "perc_iph"
MedianCI(postrain[, varname], method="exact", R = 2000)
MedianCI(negtrain[, varname], method="exact", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```



### % Lipid - IPH volume
```{r}
varname <- "perc_lipid_minus_iph"
MedianCI(postrain[, varname], method="boot", R = 2000)
MedianCI(negtrain[, varname], method="boot", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```


### IPH to Lipid volume ratio
```{r}
varname <- "iph_total_lipid_ratio"
MedianCI(postrain[, varname], method="exact", R = 2000)
MedianCI(negtrain[, varname], method="exact", R = 2000)

r <- wilcox.test(postrain[, varname], negtrain[, varname])
r

pvalues <- append(pvalues, r$p.value)
```

```{r}
round(p.adjust(pvalues, method="BH"), digits=2)
round(p.adjust(pvalues, method="BH"), digits=3)
```







